Backscatter Modulation Design for Symbiotic Radio Networks

### 2021-09-06 main.tex
:pencil: Transition probability matrix

\usepackage{adjustbox}
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{amsfonts}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{array}
\usepackage{circuitikz}
\usepackage{cite}
\usepackage{colortbl}
\usepackage{environ}
\usepackage{grffile}
\usepackage{hyperref}
\usepackage{import}
\usepackage{mathtools}
\usepackage{microtype}
\usepackage{multirow}
\usepackage{pgffor}
\usepackage{pgfplots}
\usepackage{siunitx}
\usepackage{stfloats}
\usepackage{tikz}
\usepackage{url}
\usepackage{xcolor}
\usepackage[T1]{fontenc}
\usepackage[caption=false,font=footnotesize,subrefformat=parens,labelformat=parens]{subfig}
\usepackage[short]{optidef}
% \usepackage[subtle]{savetrees}

@@

	\begin{section}{System Model}
		\begin{figure}[!t]
			\centering
			\def\svgwidth{0.9\columnwidth}
			\import{assets/}{symbiotic_radio.pdf_tex}
			\caption{A single-user single-tag symbiotic radio system.}
			\label{fi:symbiotic_radio}
		\end{figure}
		As shown in Fig.~\ref{fi:symbiotic_radio}, we propose a single-user (UE) single-tag (TG) symbiotic radio network where the RF signal generated by the single-antenna Access Point (AP) is shared by two coexisting systems. In the primary AP-UE downlink system, the AP transmits to the single-antenna user. In the secondary AP-TG-UE backscatter system, the AP acts as the carrier emitter, the user serves as the backscatter reader, and the single-antenna tag modulates its information over the reradiated RF signal by varying the reflection coefficient. Denote the AP-UE direct channel as $h_{\mathrm{D}}$, the AP-TG forward channel as $h_{\mathrm{F}}$, and the TG-UE backward channel as $h_{\mathrm{B}}$. We consider the quasi-static block fading model and assume the CSI of the direct channel and the cascaded forward-backward channel $h_{\mathrm{C}} \triangleq h_{\mathrm{B}} h_{\mathrm{F}}$ are known at the AP.\footnote{Due to the lack of RF chains at the passive tag, accurate and efficient CSI acquisition at the AP can be challenging. One possible approach is that the AP sends known pilots, the tag responds in a pre-defined manner, and the user performs least-square estimation then feeds back to the AP \cite{Bharadia2015,Yang2015b,Guo2019e}.} It is assumed that the primary symbol $s$ follows standard CSCG distribution $\mathcal{CN}(0,1)$ and the secondary symbol $c$ employs $M$-PSK modulation by \eqref{eq:backscatter_modulation}. Due to the practical constraints on switching speed and synchronization gap, the passive tag typically transmits at a much lower data rate than the AP. Hence, we assume the secondary symbol period is $N \gg 1 \in \mathbb{Z}_{++}$ times the primary symbol period and focus on the interval of one particular $c$. At symbol block $n \in \mathcal{N} \triangleq \{1,\ldots,N\}$, the user simultaneously captures the signal from both primary and secondary links as\footnote{We assume the time difference of arrival from the AP-UE path and the AP-TG-UE path are negligible compared to the symbol period \cite{Guo2019b,Liang2020,Long2020a}.}
		\begin{equation}
			y[n] = \sqrt{p} h_{\mathrm{D}} s[n] + \sqrt{\alpha p} h_{\mathrm{C}} c s[n] + w[n],
		\end{equation}
		where $p$ is the average transmit power at the AP and $w \sim \mathcal{CN}(0,\sigma^2)$ is the additive white Gaussian noise. We also define $\boldsymbol{y} \triangleq \bigl[y[1], \ldots, y[N]\bigr]^T$, $\boldsymbol{s} \triangleq \bigl[s[1], \ldots, s[N]\bigr]^T$, and $\boldsymbol{w} \triangleq \bigl[w[1], \ldots, w[N]\bigr]^T$.

@@

		To investigate how backscatter modulation potentially benefits the primary transmission, we first decode the tag symbol $c$ in presence of unknown AP symbol $\boldsymbol{s}$. Interestingly, the tag detection indeed aligns with the case of Ambient Backscatter Communications (AmBC) where the signal from the ambient source is unknown. The optimal Maximum-Likelihood (ML) detector is
		\begin{equation}
			\hat{m} = \arg \max_{m \in \mathcal{M}} p(\boldsymbol{y} \mid c_m),
		\end{equation}
		where $p(\boldsymbol{y} \mid c_m)$ denotes the probability density function of receiving $\boldsymbol{y}$ when the tag transmits $c_m$.

		During the backscatter detection, the conditional probability density function of the channel output given the tag input for $M$-PSK is
		\begin{equation}
			f_{Y \mid C}(y \mid c_m) = ?
		\end{equation}

>>>>

\usepackage{adjustbox}
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{amsfonts}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{array}
\usepackage{circuitikz}
\usepackage{cite}
\usepackage{colortbl}
\usepackage{environ}
\usepackage{grffile}
\usepackage{hyperref}
\usepackage{import}
\usepackage{mathtools}
\usepackage{microtype}
\usepackage{multirow}
\usepackage{pgffor}
\usepackage{pgfplots}
\usepackage{physics}
\usepackage{siunitx}
\usepackage{stfloats}
\usepackage{tikz}
\usepackage{url}
\usepackage{xcolor}
\usepackage[T1]{fontenc}
\usepackage[caption=false,font=footnotesize,subrefformat=parens,labelformat=parens]{subfig}
\usepackage[short]{optidef}
% \usepackage[subtle]{savetrees}

@@

	\begin{section}{System Model}
		\begin{figure}[!t]
			\centering
			\def\svgwidth{0.9\columnwidth}
			\import{assets/}{symbiotic_radio.pdf_tex}
			\caption{A single-user single-tag symbiotic radio system.}
			\label{fi:symbiotic_radio}
		\end{figure}
		As shown in Fig.~\ref{fi:symbiotic_radio}, we propose a single-user (UE) single-tag (TG) symbiotic radio network where the RF signal generated by the single-antenna Access Point (AP) is shared by two coexisting systems. In the primary AP-UE downlink system, the AP transmits to the single-antenna user. In the secondary AP-TG-UE backscatter system, the AP acts as the carrier emitter, the user serves as the backscatter reader, and the single-antenna tag modulates its information over the reradiated RF signal by varying the reflection coefficient. Denote the AP-UE direct channel as $h_{\mathrm{D}}$, the AP-TG forward channel as $h_{\mathrm{F}}$, and the TG-UE backward channel as $h_{\mathrm{B}}$. We consider the quasi-static block fading model and assume the CSI of the direct channel and the cascaded forward-backward channel $h_{\mathrm{C}} \triangleq h_{\mathrm{B}} h_{\mathrm{F}}$ are known at the AP.\footnote{Due to the lack of RF chains at the passive tag, accurate and efficient CSI acquisition at the AP can be challenging. One possible approach is that the AP sends known pilots, the tag responds in a pre-defined manner, and the user performs least-square estimation then feeds back to the AP \cite{Bharadia2015,Yang2015b,Guo2019e}.} It is assumed that the primary symbol $s$ follows standard CSCG distribution $\mathcal{CN}(0,1)$ and the secondary symbol $c$ employs $M$-PSK modulation by \eqref{eq:backscatter_modulation}. Due to the practical constraints on switching speed and synchronization gap, the passive tag typically transmits at a much lower data rate than the AP. Hence, we assume the secondary symbol period is $N \gg 1 \in \mathbb{Z}_{++}$ times the primary symbol period and focus on the interval of one particular $c$. At (primary) symbol block $n \in \mathcal{N} \triangleq \{1,\ldots,N\}$, the user simultaneously captures the signal from both primary and secondary links as\footnote{We assume the time difference of arrival from the AP-UE path and the AP-TG-UE path are negligible compared to the symbol period \cite{Guo2019b,Liang2020,Long2020a}.}
		\begin{equation}
			y[n] = \sqrt{p} h_{\mathrm{D}} s[n] + \sqrt{\alpha p} h_{\mathrm{C}} c s[n] + w[n],
		\end{equation}
		where $p$ is the average transmit power at the AP and $w \sim \mathcal{CN}(0,\sigma^2)$ is the additive white Gaussian noise. We also define $\boldsymbol{y} \triangleq \bigl[y[1], \ldots, y[N]\bigr]^T$.

@@

		To investigate how backscatter modulation potentially benefits the primary transmission, we first decode the tag symbol $c$ in presence of unknown AP symbol $\boldsymbol{s}$. Interestingly, the tag detection indeed coincides with Ambient Backscatter Communications (AmBC) where the signal from the ambient source is unknown. Denote $\mathcal{H}_m$ as the hypothesis that the tag transmits $c_m$. In such case, the received signal per symbol block follows the CSCG distribution of $y_m[n] \sim \mathcal{CN}(0,\sigma_m^2)$, where
			\sigma_m^2 \triangleq \lvert h_{\mathrm{D}} + \sqrt{\alpha} h_{\mathrm{C}} c_m \rvert^2 p + \sigma^2
		can be estimated at the user. For the ease of exposition, we sort $\{\sigma_m^2\}$ in a descending order by bijective mapping $f \colon m \mapsto i$,\footnote{Note that the mapping is not unique and the detection fails when two different constellation points yield the same energy level.} and formulate a sequence $\langle \sigma_i^2 \rangle$ where $\sigma_i^2 \ge \sigma_j^2$ for $i > j$, $i,j \in \mathcal{M}$. Following \cite{Qian2019}, the optimal Maximum-Likelihood (ML) detector boils down to the energy detector, and the likelihood ratio between hypotheses $\mathcal{H}_i$ and $\mathcal{H}_j$ is
			\Lambda_{i, j}(\boldsymbol{y}) = \frac{f(\boldsymbol{y} \mid \mathcal{H}_i)}{f(\boldsymbol{y} \mid \mathcal{H}_j)} = \left( \frac{\sigma_j^2}{\sigma_i^2} \right)^N \exp \left( \frac{(\sigma_i^2 - \sigma_j^2) z}{\sigma_i^2 \sigma_j^2} \right),
		where $f(\boldsymbol{y} \mid \mathcal{H}_i)$, $f(\boldsymbol{y} \mid \mathcal{H}_j)$ denote respectively the conditional probability density function of receiving $\boldsymbol{y}$ under hypothesis $\mathcal{H}_i$ and $\mathcal{H}_j$, and $z \triangleq \lVert y \rVert^2$ is the received signal energy over $N$ symbol blocks. The corresponding decision rule is
			\Lambda_{i, j}(\boldsymbol{y}) \underset{H_j}{\overset{H_i}{\gtrless}} 1 \iff z \underset{H_j}{\overset{H_i}{\gtrless}} T_{i, j},
		where the optimal detection threshold is
			T_{i, j} \triangleq N \frac{\sigma_i^2 \sigma_j^2}{\sigma_i^2 - \sigma_j^2} \log \frac{\sigma_i^2}{\sigma_j^2}.
		On top of this, the decision region of hypothesis $\mathcal{H}_i$ can be expressed as
		% 	\mathcal{R}(\mathcal{H}_i) \triangleq
		% 	\begin{cases}
		% 		[0, T_{1, 2}), & i = 1, \\
		% 		[T_{M-1, M}, \infty), & i = M, \\
		% 		[T_{i-1, i}, T_{i, i+1}), & \text{otherwise}
		% 	\end{cases}
			\mathcal{R}_i \triangleq [T_{i-1, i}, T_{i, i+1}),
		where we define $T_{0, 1} \triangleq 0$ and $T_{M, M+1} \triangleq \infty$. On the other hand, the transition probability matrix is
			\boldsymbol{P} =
			\begin{bmatrix}
				p_{1 \mid 1} & \ldots & p_{M \mid 1} \\
				\vdots & \ddots & \vdots \\
				p_{1 \mid M} & \ldots & p_{M \mid M}
			\end{bmatrix},
		where $p_{j \mid i}$ denotes the transition probability from channel input state $i$ to output state $j$. It can be derived as
			p_{j \mid i} = p(z \in \mathcal{R}_j \mid \mathcal{H}_i) = \int_{\mathcal{R}_j} f(z \mid \mathcal{H}_i) \dd z
			\label{eq:transition_probability}
		where $f(z \mid \mathcal{H}_i)$ is the conditional probability density function of receiving $z$ under hypothesis $\mathcal{H}_i$. Since $z$ follows chi-squared distribution with $2N$ degrees of freedom, it holds that
			f(z \mid \mathcal{H}_i) = \frac{z^{N-1} e^{-z/\sigma_i^2}}{\sigma_i^{2N} \Gamma(N)}
			\label{eq:conditional_pdf}
		where $\Gamma(\cdot)$ is the gamma function and $\Gamma(N) = (N - 1) !$ for positive integer $N$. Therefore, the transition matrix $\boldsymbol{P}$ can be calculated by substituting \eqref{eq:conditional_pdf} into \eqref{eq:transition_probability}, and the optimal input distribution can be obtained by the Blahut-Arimoto algorithm \cite{Arimoto1972,Blahut1972}.

		% During the backscatter detection, the conditional probability density function of the channel output given the tag input for $M$-PSK is
		% \begin{equation}
		% 	f_{Y \mid C}(y \mid c_m) = ?
		% \end{equation}
